---
title: "Cytopenia_VennDiagram"
author: "Usman"
date: "2023-11-02"
output: 
  html_document: 
    toc: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Prevalence of Cytopenia in Niave HIV-1 Infected Patients Initiating First-Line ART: A pilot study**


## **Introduction**
Cytopenias are common hematological manifestations in Human Immunodeficiency
Virus (HIV) infection, serving as predictors of disease progression and therapeutic
outcomes. This study aimed to determine the prevalence of cytopenias and their
association with immunosuppression in treatment-naive HIV-infected participants
undergoing highly active combined Antiretroviral Drug Therapy (cART24). A
prospective study was conducted with 44 HIV-infected individuals initiating cART,
assessing cytopenias including anemia, leucopenia, lymphocytopenia, and
thrombocytopenia, following WHO guidelines. The study employed cross-sectional and
longitudinal mixed-design two-step analysis. The median time to cART initiation was 7
days. Prevalence of cytopenia was 75% at pre-treatment baseline and 84% following
cART24. Longitudinally, anemia, leucopenia, lymphopenia, and thrombocytopenia
prevalences were 66%, 23%, 0%, and 11% at baseline, and 66%, 29%, 5%, and 20%
post-cART24, respectively. Cytopenia prevalence increased with decreasing CD4+ T
cell counts. Unicytopenia included isolated anemia (58%), leucopenia (6%), and
thrombocytopenia (6%), while 27% had bi-cytopenia and 3% pancytopenia. Anemia
and thrombocytopenia combination was most frequent. The study revealed a lack of
strong correlation between cytopenias and disease progression based on CD4+ T cell
counts. Additionally, participants with CD4+ T cell counts <200 cells/μl exhibited
decreased hemoglobin concentration following cART24. Notably, prevalence of anemia
was high in cART-naive patients, and leucopenia and neutropenia were elevated in
patients on cART, intensifying as CD4+ T cell counts decreased. These findings
emphasize the intricate hematological consequences of HIV and antiretroviral therapy,
underscoring the need for tailored preventive strategies.



```{r Cytopenia}
library(readxl)
Cyto_Prev <- read_excel("C:/Users/Usman Ola/Downloads/ERIOLUWa MAIN DATA.xlsx", 
                        sheet = "Data")
head(Cyto_Prev)
library(tidyverse)
library(ggVennDiagram)

Cyto_Prev1<- Cyto_Prev[,c(27,29,31,35)]
Cyto_Prev1
names(Cyto_Prev1)<-c("WBC","LYMP","PLAT","HGB")


Venn_Cyto<-Cyto_Prev1%>%mutate(size=WBC < 4,Plat= PLAT < 150,anem=HGB < 12)%>%
  select(anem,Plat,size)%>%
  rename(Anemia=anem,Thrombocytopenia=Plat,Leucopia=size)
Venn_Cyto
```

## **Plots**

### **Prevalenc of Cytopenia Relative to Multitude of Penia**

```{r Plots}
ggVennDiagram(lapply(Venn_Cyto, function(x)which(x==1)),
  category.names = c("A","L","T") )+
   scale_fill_gradient(low = "white",high = "white")+
    theme(legend.position = "none",plot.tag.position="bottomleft")

ggVennDiagram(lapply(Venn_Cyto, function(x)which(x==1)),
 category.names = c("Anemia","Leucopenia","Thrombocytopenia"),
  label_txtWidth = 100, set_size=4.4)+
   scale_fill_gradient(low = "white",high = "white")+
    theme(legend.position  = "bottom",plot.tag.position="bottomleft")+
     scale_x_continuous(expand = expansion(mult = .1))+
      theme(legend.position = "none",plot.tag.position="bottomleft")

```

**Figure 1** Distribution of cytopenias relative to its prevalence among treatment naïve HIV infeceted study participants at baseline and at
cART24. A : represent treatment naïve HIV infected participants while B: represent 24 weeks of weeks of cART24
